---
title: "World Politics"
subtitle: "Challenges to the Post-War Order"
author: "Michael E. Flynn"
institute: "Kansas State University"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{tikz}
  - \usepackage{pgfplots}
output:
  xaringan::moon_reader:
    css: ["default", "lecture_theme.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
      titleSlideClass: [middle, left, inverse]
---
  
```{r setup, include = FALSE, echo=FALSE, warning=FALSE}
#remotes::install_github("gadenbuie/xaringanExtra")

library(ggplot2)
library(data.table)
library(peacesciencer)
library(kableExtra)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(ggpattern)
library(ghibli)
library(ggforce)
library(patchwork)
library(scales)
library(rmapshaper)
library(sf)
library(sp)
library(tmap)
library(raster)
library(broom)
library(xaringanExtra)
library(gghighlight)
library(tidyquant)
library(flynnprojects)

xaringanExtra::use_scribble()
xaringanExtra::use_editable(expires = 365)

knitr::opts_chunk$set(comment = '', dpi = 400, fig.width = 8, fig.height = 5, warning = FALSE, echo = FALSE, message = FALSE)

sysfonts::font_add_google("Oswald", family = "oswald")
showtext::showtext_auto()

update_geom_defaults("text", list(size = 3, family = "oswald", fontface = "plain", lineheight = 1))


```


# Lecture Overview

1. Background Review

2. Emerging Challenges

3. Moving Forward



---
class: inverse, center, middle

# Background Review




---
# Background Review

Four primary "pillars" of the international order

1. Sovereignty

2. Collective Security

3. Bretton Woods System

4. Promotion of Human Rights and Democracy

--

These things are based on and implemented using

- Liberal values

- Multilateral cooperation

- US Hegemony


---
# Background Review

Key Institutions

- United Nations

- General Agreement on Tariffs and Trade (World Trade Organization)

- International Monetary Fund

- World Bank

- North Atlantic Treaty Organization (NATO)




---
class: center, middle

```{r, dpi=400}

bases <- troopdata::basedata

map <- rnaturalearth::ne_countries(returnclass = "sf") %>% filter(name != "Antarctica")

ggplot() +
  geom_sf(data = map, aes(group = name, geometry = geometry)) +
  geom_point(data = bases, aes(x = lon, y = lat), color = "darkblue", alpha = 0.7) +
  theme_void() +
  theme(text = element_text(family = "oswald", face = "bold", size = 20)) +
  labs(title = "US military facilities around the world, 1950-2015")


```



---
class: center, middle

```{r}

data <- readxl::read_xlsx("~/Dropbox/Data Files/IMF/Table_1_World_Currency_Composition_.xlsx") |> 
  dplyr::rename("Currency" = 1) |> 
  filter(grepl(".*Shares of.*", Currency)) |> 
  slice(-1) |> 
  dplyr::select(-c(2, 3, 4, 5)) |> 
  filter(!grepl(".*Unallocated.*", Currency)) |> 
  mutate(Currency = gsub("Shares of ", "", Currency),
         Currency = str_to_title(Currency),
         Currency = ifelse(Currency == "U.s. Dollars", "U.S. Dollars", Currency)) |> 
  pivot_longer(cols = 2:97) |> 
  mutate(name = gsub("...", "", name),
         value = round(as.numeric(value), 2),
         name = as.numeric(name)) |> 
  group_by(Currency) |> 
  dplyr::mutate(year = rep(1999:2022, each = 4),
                quarter = rep(c(1, 2, 3, 4), length(Currency)/4),
                time = as.numeric(glue::glue("{year}.{quarter}"))) |> 
  drop_na(everything()) |> 
  tidyr::fill(value, 
              .direction = c("downup")) 


ggplot(data, aes(x = year, y = value, fill = Currency, group = Currency)) +
  geom_area() +
  theme_flynn(base_family = "oswald", base_size = 11) +
  viridis::scale_fill_viridis( discrete = TRUE) +
  scale_y_continuous(breaks = seq(0, 100, 20), limits = c(0, 101)) +
  labs(x = "Year",
       y = "Percent",
       title = "Currency shares of global foreign exchange reserves, 1999-2022",
       subtitle = "The US Dollar has long been the dominant reserve currency",
       caption = "Data available from International Monetary Fund, Currency Composition of Foreign Exchange Reserves \n http://data.imf.org/?sk=E6A5F467-C14B-4AA8-9F6D-5A09EC4E62A4")


```




---
class: center, middle, inverse

# Emerging Challenges



---
class: left, top

# Emerging Challenges


What are the issues?

- (Bad) Non-state actors

- Nuclear proliferation

- Biological weapons

- Cyberattacks

- Revisionist powers (e.g. China, Russia, etc.)

- Global health and pandemics

- Climate change and environmental degradation 

- Populist movements 

- Political and economic instability



---
class: center, middle

```{r}

nuclearpowers <- c("United States of America", "United Kingdom", "Russia", "China", "France", "Pakistan", "India", "Dem. Rep. Korea", "Israel")

map <- rnaturalearth::ne_countries(returnclass = "sf") %>% 
  mutate(nuclear = ifelse(name %in% nuclearpowers, "Yes", "No")) %>% 
  filter(name != "Antarctica")

ggplot() +
  geom_sf(data = map, aes(geometry = geometry, fill = nuclear), color = "black", size = .1) +
  viridis::scale_fill_viridis(discrete = TRUE, begin = 0.25, end = 0.85) +
  theme_void() +
  theme(text = element_text(family = "oswald", face = "bold", size = 20)) +
  labs(title = "States possessing nuclear weapons",
       fill = "Nuclear Weapons")


```




---
class: center, middle

```{r}

gdp <- peacesciencer::create_stateyears(system = "gw") %>% 
  add_sdp_gdp() %>% 
  filter(year >= 1950) %>% 
  dplyr::rename("ccode" = "gwcode")

capabilities <- peacesciencer::cow_nmc %>% 
  filter(year >= 1950) %>% 
  left_join(gdp, by = c("ccode", "year"))


ggplot(capabilities %>% filter(year == 2015), aes(x = exp(wbgdp2011est), y = cinc)) +
  geom_point(alpha = 0.75, shape = 21, fill = "orange", size = 3) +
  ggrepel::geom_label_repel(aes(label = statename), size = 3) +
  theme_flynn(base_family = "oswald", base_size = 11) +
  scale_x_continuous(labels = scales::dollar_format()) +
  labs(x = "World Bank GDP Estimate",
       y = "CINC Score",
       title = "GDP and Correlates of War CINC Scores, 2015",
       caption = "Data obtained from the {peacesciencer} package.")

```



---
class: center, middle

```{r}


ggplot(capabilities %>% filter(year == 2015), aes(x = exp(wbgdp2011est), y = milex)) +
  geom_point(alpha = 0.75, shape = 21, fill = "orange", size = 3) +
  ggrepel::geom_label_repel(aes(label = statename), size = 3) +
  theme_flynn(base_family = "oswald", base_size = 11) +
  scale_x_continuous(labels = scales::dollar_format()) +
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(x = "World Bank GDP Estimate",
       y = "Military Expenditures (thousands of USD)",
       title = "GDP and military expenditures, 2015",
       caption = "Data obtained from the {peacesciencer} package.")
```



---
class: center, middle, inverse

# Moving Forward



---
class: left, top

# Moving Forward

So what's to be done?

- Lots of different issues, but there are also lots of common themes and characteristics

- The work is increasingly interconnected, and so externalities from human behavior (both positive and negative) are increasingly visible

- People have shared interests in solving these problems

- Most human interaction is, in fact, benign/cooperative

  
  
---
class: left, top

# Moving Forward

So what are some of these common themes/frameworks?

  - Collective action 
  
  - Cooperation and coordination (think about rules and incentive structures)

  - Bargaining (think about people's interests and what's possible)
  
--

Factors that shape or affect the specifics of these frameworks

  - Actors: Who are the relevant players?
  
  - Interests: What do the players want? What do they value?
  
  - Institutions: What are the prevailing rules (formal or informal) that govern actors' behavior?



---
# Moving Forward

Other tips:

- Most of the things we care about are caused by lots of other different things (complex systems)...

- ...But not everything has the same amount of causal influence/importance

- Most of the causal relationships that we care about are probabilistic—not deterministic (i.e. more or less likely)


